OpenMPIR

### 2017-09-01 paper.tex
Fixed typo on correctness
Added example codegen

\usepackage{booktabs} % For formal tables
\usepackage{multicol}
\usepackage{graphicx}

@@

While we did implement the codegen for the \texttt{task\_spawn} and
\texttt{task\_wait} constructs, it's worth noting here that even for these
pragmas, the implementation is incomplete. Currently, any clauses modifying the 
behaviour of the pragmas is ignored. Probably the most common semantics this
will change of variable semantics, e.g. \texttt{shared} vs. \texttt{private}.
Surprisingly, this had very effect on the correctness of the entire Barcelona
OpenMP task suite. Indeed, it is likely that fixing this issue would increase
the performance of this work marginally, due to reducing the number of memory
copies for variables declared \texttt{private} in OpenMP code.

@@

@@

To better understand how the OpenMP to Tapir compiler works, we turn to the
\texttt{fib} example from Section~\ref{Sec:OpenMP}. In Figure~\ref{Fig:Example}
we show how the code generated differs between the original Clang
implementation, which generates vanilla LLVM with calls into the OpenMP
runtime, and.

\section{Evaluation}

For evaluation we compared performance on the Barcelona OpenMP task suite to 
existing OpenMP implementations. The Barcelona OpenMP task suite is a set of 
tests intended to test performance of OpenMP implementations using both
irregular and regular tasking. The benchmarks are described in
Figure~\ref{Fig:benchmark_desc}.

@@

In particular, we compare to GCC 7.1, Clang 4.0.1, and Intel 17.0.0. The
machine used is a two socket Intel Xeon E5-2693 v3 machine with 132GB of
memory, running Linux 4.11.4. It's worth noting that each of the other 
implementations comes with it's own runtime. In this sense, performance is
a function of both the compiler and any optimizations it's able to perform, 
and the runtime. Understanding the interaction of these two parts is non-trivial,
as we will see when discussing results.

@@

\begin{itemize}
\item Non-determinism in result correctness
\item Segmentation faults
\item Timeouts (10 Minutes)
\end{itemize}

We set a timeout of 10 minutes, as at least one implementation was always
finishing within ~10 seconds, anything running more than 60 times slower
becomes irrelevant. DUe to time constraints, correctness checks were not
performed on every run, so it is possible some non-determinism was missed.

@@

@@

Another import consideration is that we can break reasons for performance
discrepencies into a few categories:

>>>>

\usepackage{booktabs} % For formal tables
\usepackage{multicol}
\graphicspath{{./figs/}}

@@

While we did implement the codegen for the \texttt{task\_spawn} and
\texttt{task\_wait} constructs, it's worth noting here that even for these
pragmas, the implementation is incomplete. Currently, any clauses modifying the 
behaviour of the pragmas is ignored. Probably the most common semantics this
will change of variable semantics, e.g. \texttt{shared} vs. \texttt{private}.
Surprisingly, this had little effect on the correctness of the entire Barcelona
OpenMP task suite. Indeed, it is likely that fixing this issue would increase
the performance of this work marginally, due to reducing the number of memory
copies for variables declared \texttt{private} in OpenMP code.

@@

All code used for the implementation is available at
\texttt{https://github.com/lanl/openmpir}. The version used for this paper is
tagged as LLVM17.

@@

To better understand how the OpenMP to Tapir compiler works, we turn to the
\texttt{fib} example from Section~\ref{Sec:OpenMP}. In Figure~\ref{Fig:Example}
we show what Tapir code is generated.

\begin{figure*}
\begin{tabular}{| c c c | c c c |}
\hline
 & 
\begin{lstlisting}
int fib(int n) 
{                    // A
  if (n < 2)
    return n;
  else
  {
    int x, y;
    #pragma omp task
      x = fib(n-1);  // B
    #pragma omp task
      y = fib(n-2);  // C
    #pragma omp taskwait
    return (x+y);    // D
  }
}
\end{lstlisting}
& & &
\begin{lstlisting}
if.end:                                 
  detach within %syncreg, label %det.achd, label %det.cont

det.achd:                                 
  %2 = load i32, i32* %n.addr, align 4
  %sub = sub nsw i32 %2, 1
  %call = call i32 @fib(i32 %sub)
  store i32 %call, i32* %x, align 4
  reattach within %syncreg, label %det.cont

det.cont:                                   
  detach within %syncreg, label %det.achd1, label %det.cont4

det.achd1:                                
  %3 = load i32, i32* %n.addr, align 4
  %sub2 = sub nsw i32 %3, 2
  %call3 = call i32 @fib(i32 %sub2)
  store i32 %call3, i32* %y, align 4
  reattach within %syncreg, label %det.cont4

det.cont4:                                  
  sync within %syncreg, label %sync.continue
\end{lstlisting}
 &  \\
\hline
\end{tabular}

\caption{Compilation to Tapir}
\label{Fig:Example}
\end{figure*}

\section{Evaluation} \label{Sec:Evaluation}

For evaluation we compared performance on the Barcelona OpenMP task suite to 
existing OpenMP implementations. The Barcelona OpenMP task suite is a set of 
tests intended to test performance of OpenMP implementations using both
irregular and regular tasking. The benchmarks are described in
Figure~\ref{Fig:benchmark_desc}. The benchmark suite has added an unbalanced
tree search benchmark since the paper was published.  Unfortunately, this seems
to be buggy, and failed on every implementation for even medium input sizes. We
have therefore left it out of our evaluation.

@@

There are two classes of input sizes for the benchmarks. Some, like Floorplan,
require input files. For each of these, we chose the largest provided. For the
benchmarks with a parameter to adjust the size of the input, we attempted to 
choose a parameter size so that the fastest implementation was on the order of
seconds. The exact inputs used are:

\begin{itemize}
\item \texttt{Alignment: -f prot.100.aa} 
\item \texttt{FFT: -n 335544320}
\item \texttt{Fibonacci: -n 48}
\item \texttt{Floorplan: -f input.20}
\item \texttt{Health: -f large.input}
\item \texttt{NQueens: -n 15}
\item \texttt{Sort -n 335544320}
\item \texttt{SparseLU: -n 100 -m 100}
\item \texttt{Strassen: -n 8192}
\end{itemize}

For other implementations, we compare to GCC 7.1, Clang 4.0.1, and Intel
17.0.0. The machine used is a two socket Intel Xeon E5-2693 v3 machine with
132GB of memory, running Linux 4.11.4. All benchmarks were run using all 64
hyper-threads. 32 thread tests were done as well and showed qualitatively
similar results. It's worth noting that each of the other implementations comes
with it's own runtime. In this sense, performance is a function of both the
compiler and any optimizations it's able to perform, and the runtime.
Understanding the interaction of these two parts is non-trivial, as we will see
when discussing results.

@@

\begin{itemize}
\item Segmentation faults
\item Timeouts (10 Minutes)
\item Non-determinism in result correctness
\end{itemize}

We set a timeout of 10 minutes, as at least one implementation was always
finishing within ~10 seconds, anything running more than 60 times slower
becomes irrelevant. Due to time constraints, correctness checks were not
performed on every run, so it is possible some non-determinism was missed.

Each variant was run 10 times, with the height of the bars representing the
mean and the error bars representing standard deviation. In cases where there
was timeout or segmentation fault, no time is reported, and the run is marked
as faulty.

@@

Putting failures aside for the moment, performance for our implementation is
quite strong, generally outperforming or matching the best of existing
implementations. For benchmarks where the overhead to work ratio is high is 
where the Tapir implementation really shines. For example, for the
\texttt{Fibonacci} benchmark, our implementation finishes significantly faster
than any of the others. While ICC failed for that sized input, we did test with
smaller inputs and found its similarly outpaced by the Tapir implementation.
\texttt{NQueens} and \texttt{FFT} show similar behaviour, to lesser degrees. In 
contrast, benchmarks that have a lower overhead to work ratio unsurprisingly 
differ less in performance. For example,

@@

Another import consideration is that we can break reasons for performance
discrepancies into a few categories:
